---
title: "exercise-8"
author: "Minsoo Joo"
date: "2025-03-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 1

```{r}
library(tidyverse)
library(skimr)
```

```{r}
f<-"https://raw.githubusercontent.com/difiore/ada-datasets/main/Street_et_al_2017.csv"
d<-read_csv(f, col_names=TRUE)

skim(d)
```
# Step 2

```{r}
plot_1 <- d %>% 
  ggplot(aes(x=Group_size, y=ECV)) +
  geom_point(shape = 21, colour = "red", fill = "black", size=1, na.rm = TRUE) +
  ggtitle("ECV vs Group Size") +
  theme_minimal()
plot_2 <- d %>% 
  ggplot(aes(x=Longevity, y=ECV)) +
  geom_point(shape = 21, colour = "red", fill = "black", size=1, na.rm = TRUE) +
  ggtitle("ECV vs Longevity") +
  theme_minimal()
plot_3 <- d %>% 
  ggplot(aes(x=Weaning, y=ECV)) +
  geom_point(shape = 21, colour = "red", fill = "black", size=1, na.rm = TRUE) +
  ggtitle("ECV vs Weaning") +
  theme_minimal()
plot_4 <- d %>% 
  ggplot(aes(x=Repro_lifespan, y=ECV)) +
  geom_point(shape = 21, colour = "red", fill = "black", size=1, na.rm = TRUE) +
  ggtitle("ECV vs Repro_lifespan") +
  theme_minimal()
library(cowplot)
plot_grid(plot_1, plot_2, plot_3, plot_4)
```

# Step 3
```{r}
d_clean <- d %>% 
  filter(!is.na(ECV) & !is.na(Group_size))
(beta_1 <- cov(d_clean$Group_size, d_clean$ECV)/var(d_clean$Group_size))
(beta_0 <- mean(d_clean$ECV) - beta_1*mean(d_clean$Group_size))
```

# Step 4
I got the same results.
```{r}
m <- lm(ECV~Group_size, data=d_clean)
summary(m)
```

# Step 5
Yes, the regression coefficients differ among groups.
The slope is highest for Platyrrhini (1.965), followed by Strepsirhini (1.841), and lowest for Catarrhini (1.146).
Group size has marginal significant relationship with Catarrhini, while has high significant relationships with Platyrrhini and Strepsirrhini.
To determine whether the differences are statistically significant, ANOVA or ANCOVA can be used.
```{r}
d_Catarrhini <- d %>% 
  filter(Taxonomic_group=='Catarrhini') %>% 
  filter(!is.na(ECV) & !is.na(Group_size))
d_Platyrrhini <- d %>% 
  filter(Taxonomic_group=='Platyrrhini') %>% 
  filter(!is.na(ECV) & !is.na(Group_size))
d_Strepsirhini <- d %>% 
  filter(Taxonomic_group=='Strepsirhini') %>% 
  filter(!is.na(ECV) & !is.na(Group_size))

m1 <- lm(ECV~Group_size, data=d_Catarrhini)
m2 <- lm(ECV~Group_size, data=d_Platyrrhini)
m3 <- lm(ECV~Group_size, data=d_Strepsirhini)

summary(m1)
summary(m2)
summary(m3)
```

# Step 6
I included comments in the code.
```{r}
SSY <- sum((m$model$ECV - mean(m$model$ECV))^2) # y - y_mean
SSR <- sum((m$fitted.values - mean(m$model$ECV))^2) # y_predicted - y_mean
SSE <- sum((m$model$ECV - m$fitted.values)^2) # or SSE <- sum(m$residuals^2) -> y - y_predicted
SSY == SSE + SSR # for checking
(df_regression <- 1)  # p = 1 (the number of predictor variables)
(df_error <- nrow(d_clean) - df_regression - 1)  # n - p - 1
(df_y <- nrow(d_clean) - df_regression)  # n - p

MSR <- SSR/df_regression  # mean variance explained by the regression equation
MSE <- SSE/df_error  # mean remaining variance
MSY <- SSY/df_y  # mean overall variance

SSX <- sum((m$model$Group_size - mean(m$model$Group_size))^2)  # how much x variation there is
(SEbeta1 <- sqrt(MSE/SSX)) # the standard error for the slope coefficient!!!

beta_1 + qt(p = c(0.025, 0.975), df=df_error)*SEbeta1 # 95% of CI!!!

t.calc <- beta_1 / SEbeta1 # t-statistic
(p.calc = 2 * (1 - pt(abs(t.calc), df = nrow(m$model) - 2))) # p-value!!!!

summary(m) # same results
confint(m) # same results
```

# Step 7
```{r}
m <- lm(ECV ~ Group_size, data = d_clean)
beta1 <- coef(m)[2] # Slope coefficient

n_permutations <- 1000
perm_beta1 <- numeric(n_permutations) # Store permuted slope values

# permutation approach to calculate p value
# We should permute the dependent variable (ECV).
for (i in 1:n_permutations) {
  permuted_ECV <- sample(d_clean$ECV)
  perm_model <- lm(permuted_ECV ~ d_clean$Group_size)
  perm_beta1[i] <- coef(perm_model)[2]
}

SE_permutation <- sd(perm_beta1) # standard error
t_stat_perm <- beta1 / SE_permutation # t-statistic
p_value <- 2 * (1 - pt(abs(t_stat_perm), df = nrow(d_clean) - 2)) # Calculate p-value (theory-based method)

p_value
```

# Step 8
The slope coefficient is different from zero, indicating that the relationship is likely to be significant.
```{r}
m <- lm(ECV ~ Group_size, data = d_clean)
beta1 <- coef(m)[2]

n_bootstrap <- 1000
bootstrap_beta1 <- numeric(n_bootstrap)

for (i in 1:n_bootstrap) {
  boot_sample <- d_clean[sample(nrow(d_clean), replace = TRUE), ]
  boot_model <- lm(ECV ~ Group_size, data = boot_sample)
  bootstrap_beta1[i] <- coef(boot_model)[2]
}
#quantile method
CI_quantile <- quantile(bootstrap_beta1, c(0.025, 0.975))
CI_quantile

#theory-based method
SE_bootstrap <- sd(bootstrap_beta1)
t_critical <- qt(0.975, df = nrow(d_clean) - 2)
CI_theory <- beta1 + c(-1, 1) * t_critical * SE_bootstrap
CI_theory
```


